% Copyright (c)  2005  Bram de Greve.
%
% Permission is granted to copy, distribute and/or modify
% this document under the terms of the GNU Free Documentation
% License, Version 1.2 or any later version published by the
% Free Software Foundation; with the Invariant Sections being
% the chapter ``Introduction", no Front-Cover Texts, and no
% Back-Cover Texts.  A copy of the license is included in the
% section entitled "GNU Free Documentation License".

\chapter{Shaders module}

\kw{shaders} is the module that is all about defining the appearance of the geometry objects.

\section{surface shaders}

\subsection{Unshaded}\label{shaders.Unshaded}

The most simple shader of all.  It simply returns the texture value

\subsubsection*{constructors}
\begin{itemize}
	\item \kw{Unshaded()}:
		creates a shader with a constant white texture
	\item \kw{Unshaded(colour)}:
		creates a shader with colour texture.
\end{itemize}

\subsubsection*{members}
\begin{itemize}
	\item \kw{colour}:
		get and set the colour texture.
\end{itemize}



\subsection{Lambert}\label{shaders.Lambert}

Models a perfect Lambertian surface.

\subsubsection*{constructors}
\begin{itemize}
	\item \kw{Lambert()}:
		creates a shader with a constant white texture
	\item \kw{Lambert(diffuse)}:
		creates a shader with a diffuse colour texture.
\end{itemize}

\subsubsection*{members}
\begin{itemize}
	\item \kw{diffuse}:
		get and set the diffuse colour texture.
\end{itemize}



\subsection{Simple}\label{shaders.Simple}

Models a classical surface material

\subsubsection*{constructors}
\begin{itemize}
	\item \kw{Simple()}:
		creates a shader with a constant white texture
	\item \kw{Simple(diffuse)}:
		creates a shader with a diffuse colour texture.
	\item \kw{Simple(diffuse, specular)}:
		creates a shader with a diffuse and specular colour texture.
\end{itemize}

\subsubsection*{members}
\begin{itemize}
	\item \kw{diffuse}:
		get and set the diffuse colour texture.
	\item \kw{specular}:
		get and set the diffuse colour texture.
	\item \kw{specularPower}:
		get and set the sharpness texture for the specular lighting.  This texture is evaulated as a scalar.
	\item \kw{reflectance}:
		get and set the reflectance colour texture.
	\item \kw{transmittance}:
		get and set the transmittance colour texture.
	\item \kw{refractionIndex}:
		get and set the texture that gives the relative index of refraction of the interior material with respect to the exterior metarial.  This texture is evaluated as a scalar.
\end{itemize}

\section{volume shaders}

\subsection{ExponentialFog}\label{shaders.ExponentialFog}

\begin{equation}
	\sigma \left( h \right) = a \exp \left( -bh \right)
\end{equation}

\begin{equation}
	\begin{array}{lcl}
		\sigma \left( t \right) &=& a \exp \left( -b \left( \Vect{op} \cdot \Vect u + t \, \Vect \omega \cdot \Vect u \right) \right)\\
		&=& \frac A B \exp \left( -B t ) \right)
			\text{ with } \begin{cases}
				A = a \exp \left( -b \, \Vect{op} \cdot \Vect u \right)\\
				B = b \, \Vect \omega \cdot \Vect u
			\end{cases}
	\end{array}
\end{equation}

\begin{eqnarray}
	\tau \left( d \right) &=& \int_0^d \sigma \left( t \right) \D t = \frac A B \left( 1 - \exp \left( -B t \right) \right)\\
	\tau \left( d \right) &\approx& A d \left(1 - \frac{B d} 2\right) \Leftrightarrow \Abs{B d} \ll 1
\end{eqnarray}

\begin{equation}
	\alpha_r \left( d \right) = 1 - \exp \left( -\tau \left( d \right) \right)
\end{equation}

\begin{equation}
	\alpha_{r,max} = \lim_{d \rightarrow \infty} \alpha_r \left( d \right) =
		\begin{cases}
			1 - \exp \left( -\frac A B \right) & \Leftrightarrow B > 0\\
			1 & \Leftrightarrow B \leq 0
		\end{cases}
\end{equation}

\begin{equation}
	P \left(d\right) = \Prob{x < d} = \frac {\alpha_r \left( d \right)} {\alpha_{r,max}}
\end{equation}

\begin{equation}
	P^{-1} \left( \xi \right) = \begin{cases}
		- \frac 1 B \ln \left(1 + \frac B A \ln \left(1 - \alpha_{r,max} \xi\right)\right) & \Leftrightarrow B \neq 0\\
		- \frac 1 A \ln \left(1 - \alpha_{r,max} \xi\right) & \Leftrightarrow B = 0
	\end{cases}
\end{equation}

\begin{equation}
	p \left(d\right) = \frac A {\alpha_{r,max}} \exp \left( -\tau \left(d\right) - B d \right)
\end{equation}
